---
title: "project report"
output: html_document
---

First download the files and load relevant library
```{r}
library(caret)
library(rpart)
library(randomForest)

set.seed(12345)
trainingfile<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingfile<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
df<-read.csv(trainingfile,header = T,na.strings = c("","NA","#DIV/0!"))
```


Partition the dataset into training and testing set. 
```{r}
inTrain<-createDataPartition(df$classe,p=0.8,list=F)
training<-df[inTrain,]
testing<-df[-inTrain,]
```
Quick look at the data
```{r}
summary(training)
```

There are many colnums with na values or values that are not relevant. Remove columns with na values and columnes that are not relevant. 

```{r}
nafilter<-!(colSums(is.na(training))>0) #& !colSums(training=="NA")>0

training<-training[,nafilter]
training<-training[-c(1,3,4,5)]

testing<-testing[,nafilter]
testing<-testing[-c(1,3,4,5)]
```

look at the dimmision with the data
```{r}
dim(training)
```

There are only 56 columnes reserved. 


Firstly try a decion tree model, which hopefully will give some insights.

```{r}

model1<-rpart(classe~., data=training,method="class")
rtrain<-predict(model1,newdata=training,type="class")
confusionMatrix(rtrain,training$classe)
rtest<-predict(model1,newdata=testing,type="class")
confusionMatrix(rtest,testing$classe)
```
The results are pretty ugly even on the training set. try with a more complicated random forrest model.  
```{r}
#model2<-train(classe~., method="rf",data=training,tcControl=tc)
print("random forest result#########")
model2<-randomForest(classe~., data=training,ntree=2000)
confusionMatrix(predict(model2,newdata=training),training$classe)
confusionMatrix(predict(model2,newdata=testing),testing$classe)
```

results look promising and because it is random forrest, the expected out of sample result should be quite similar. load test data and make predicion. 
```{r}
va<-read.csv(testingfile,header=T)
levels(va$new_window) <- levels(training$new_window)
predictions<-predict(model2,newdata=va)

answers<-as.character(predictions)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```
pml_write_files(answers)




